\chapter{Introduction}

\lignum\ is a tree model developed by the researches in the  Finnish
Forest Research Institute, The University of Helsinki and The Helsinki
University of  Technology.   The model \lignum\  has  been designed to
capture the three dimensional structure of  a tree species and, at the
same time, to be able to model its growth and development.

The three dimensional structure of a tree is captured with the help of
simple structural  units close to their  real  world counterparts. The
metabolic  functioning (photo  synthesis,  respiration,  transport  of
nutrients) is modelled by  associating appropriate functions to  these
structural units.  The growth and  senescence is  modelled by changing
the size of  these units, creating  new units and deleting these units
during a simulation.

The  software  that implements \lignum\  has   been programmed in C++.
Since   \lignum\   was      first    introduced  \cite{salminen:omt94,
perttunen:aob96}  a lot  of things   have  happened. During the  first
implementations of the model  the terminology was not  established and
this inevitably was reflected in the software. Also, the main tool the
C++ programming  language  was still  under development.   Thirdly, as
characterizing any  research project many things  were found  by trial
and  error, by experimenting    different alternatives. This  too, was
reflected in the software.

Today many things are different.   The terminology used in \lignum\ is
now established and approved by the scientific community. The C++ ANSI
Committee has finished  its work.  The major addition  to the language
from the  point of view of  \lignum\ is the  Standard Template Library
(STL)  now  providing  the  ubiquitous abstract  datatypes  needed  in
software development.   The research  work is still  going on,  but we
have now  established a solid  set of parameters, state  variables and
functions that can be used to describe a functioning of a tree with
\lignum. In addition, \lignum\ is no longer a research project \it per
se\rm, but also used by, e.g.,  the students of Helsinki University of
Technology.

All this leads to only one  conclusion: The time has  come for a major
rewrite of the software to keep The Tower of Babel from rising.

This document  describes   the C++  implementation of the   tree model
\linebreak \lignum.  The  classes, algorithms and  other utilities are
available in the library called libLGM.a and related C++ header files.
The  implementation uses   the STL  designed   as  part  of  ANSI  C++
specification. In fact, if you understand  the principles of STL-library
it  should  be relatively  easy for you  to  undestand  the design and
implementation of \lignum.

The software implementing \lignum\ is a result  of an ongoing research
work,  thus only  a   momentary  (a ``snapshot'') description   can be
given. The interested  \lignum\ user  is  encouraged to  report errors
(both in the software and in the documentation) to the authors.


\chapter{Programming Style}
Please read the section Programming Style in the document c++adt.

