\section{Getting Started with LIGNUM}

\subsection{The Units of LIGNUM} 

The  main design  principle in  \lignum is  to model  a tree  with few
simple  basic units that  correspond to  the organs  of the  tree. The
units  must  be  such that  they  allow  a  detailed analysis  of  the
structure and the development of the tree during its life time and and
still keeping  the model  computationally manageable to  create forest
stands consisting of different  tree species. Ideally the units should
be such that they can  be divided into more detailed tree compartments
if necessary.

Currently a  tree is modelled in  \lignum with five  units.  The three
basic units are  tree segment, branching point and  bud and the fourth
is an  aggregate of these basic  units called axis. The  fifth unit is
naturally the tree, an aggregate of axises.

The tree segment consists of  sapwood, heartwood, bark and foliage and
it captures the main metabolic  functiong in a tree.  The tree segment
corresponds roughly  to the annual  shoot but in the  strict modelling
sense  this  is  not  necessarily  always the  case,  especially  when
modelling heartwood trees.  

The bud  is  an embryonic shoot, the  growing  point of  the tree from
which new shoots, leaves and flowers may develop.  Buds can be further
divided into  terminal buds located at the  top of axises  and lateral
(or axillary) buds  located in the  tree segment.  More  precisely for
heartwood trees  lateral buds  are located where  the leaf  petiole is
attached to a tree segment.

The branching point is the place  where two or  more tree segments are
connected  to  each other.  The  aggregate unit  axis corresponds to a
stem or a branch of a tree.

\subsection{Design of the Program Architecture}

The  design of  any  computer  program includes  the  decision how  to
represent information  and concepts from the real  world. C++ supports
many   programming   methodologies   (perhaps  most   notably   object
orientation)  but  the  emergence  of the  Standard  Template  Library
emphasizes   generic   programming   with  abstract   datatypes   that
encapsulates data and algorithms  working with this data. Two language

constructs  in C++  support this  approach: classes  and  templates. A
class  defines a  datatype consisting  a  set of  possible states  and
operations  defining  transitions between  those  states. Template  is
simply C++ term for a parameterized type.

As a tree  is represented with five units  in \lignum a natural choice
is to represent each unit as a class each with a set of attributes and
operations. Further more as explained   in publications these  classes
are  organized   in a  hierarchy  with  a  common  base   class called
TreeCompartment.  This design allows us  to utilize e.g., polymorphism
when implementing metabolic processes.  

However, one of the problems with the  software used to simulate young
Scots pine trees, Jack pines, sugar maples etc. is  that it has proved
to   be difficult to modify.  Typically   when creating and validating
tree  models   with \lignum  one  needs to   experiment with parameter
values, different light models, metabolic processes etc.

To be continued....

